---
title: "Baseline SIRS Simulation Run"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
find_project_root <- function(start = getwd()) {
  current <- normalizePath(start, mustWork = TRUE)
  while (!file.exists(file.path(current, "DESCRIPTION")) || !dir.exists(file.path(current, "R"))) {
    parent <- dirname(current)
    if (identical(parent, current)) {
      stop("Could not locate the SIRSsim project root.", call. = FALSE)
    }
    current <- parent
  }
  current
}
project_root <- find_project_root()
knitr::opts_knit$set(root.dir = project_root)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)

analysis_dir <- file.path(project_root, "analysis", "survival_baseline")
output_dir <- file.path(analysis_dir, "outputs")
raw_dir <- file.path(output_dir, "raw")
derived_dir <- file.path(output_dir, "derived")
figure_dir <- file.path(output_dir, "figures")

embed_png <- function(path, alt) {
  sprintf("![%s](%s)", alt, knitr::image_uri(normalizePath(path, mustWork = TRUE)))
}
```

## Purpose

This report documents the exact baseline SIRS simulation used for the survival
analysis. The simulation uses a `10 x 10` matrix with the center cell initially
infected, SIRS transitions, and mortality enabled.

## Parameters

```{r parameters}
parameters <- data.frame(
  parameter = c(
    "rows",
    "columns",
    "initial_infection",
    "model",
    "prob_infect",
    "imm_prob",
    "allow_death",
    "fat_prob",
    "seed",
    "full_log"
  ),
  value = c("10", "10", "center seeded", "SIRS", "0.25", "0.70", "TRUE", "0.15", "94128", "TRUE")
)

parameters
```

## Raw Outputs

```{r load-outputs}
full_log <- read_csv(file.path(raw_dir, "full_log.csv"), show_col_types = FALSE)
history <- read_csv(file.path(raw_dir, "history.csv"), show_col_types = FALSE)
final_matrix <- read_csv(file.path(raw_dir, "final_matrix.csv"), show_col_types = FALSE)
survival_endpoints <- read_csv(file.path(derived_dir, "survival_endpoints.csv"), show_col_types = FALSE)

list(
  full_log_rows = nrow(full_log),
  full_log_columns = names(full_log),
  simulation_steps = max(history$step),
  endpoint_rows = nrow(survival_endpoints)
)
```

## Simulation History

```{r history-table}
history
```

```{r history-plot, echo=FALSE, results='asis'}
cat(embed_png(file.path(figure_dir, "simulation_history.png"), "Simulation history"))
```

## Final Matrix

```{r final-matrix-table}
final_matrix
```

```{r final-matrix-plot, echo=FALSE, results='asis'}
cat(embed_png(file.path(figure_dir, "final_matrix.png"), "Final matrix state"))
```

## Full Log Structure

The full log contains one row per cell per step, including step `0` before any
state transitions.

```{r full-log-preview}
head(full_log, 12)
```

```{r full-log-validation}
expected_rows <- 100 * (max(history$step) + 1)
validation <- data.frame(
  check = c(
    "full_log contains expected rows",
    "survival_endpoints has one row per cell",
    "event indicators are logical"
  ),
  passed = c(
    nrow(full_log) == expected_rows,
    nrow(survival_endpoints) == 100 && length(unique(survival_endpoints$cell_id)) == 100,
    all(vapply(survival_endpoints[c("event_infection", "event_recovery", "event_death")], is.logical, logical(1)))
  )
)

validation
```

The survival reports use the standard `R` survival convention shown in
`Surv(time, status)`: event indicators are `TRUE` for observed events and
`FALSE` for censored observations.
