---
title: "Baseline SIRS Survival Analysis"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
find_project_root <- function(start = getwd()) {
  current <- normalizePath(start, mustWork = TRUE)
  while (!file.exists(file.path(current, "DESCRIPTION")) || !dir.exists(file.path(current, "R"))) {
    parent <- dirname(current)
    if (identical(parent, current)) {
      stop("Could not locate the SIRSsim project root.", call. = FALSE)
    }
    current <- parent
  }
  current
}
project_root <- find_project_root()
knitr::opts_knit$set(root.dir = project_root)
library(survival)
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(broom)

analysis_dir <- file.path(project_root, "analysis", "survival_baseline")
output_dir <- file.path(analysis_dir, "outputs")
derived_dir <- file.path(output_dir, "derived")
figure_dir <- file.path(output_dir, "figures")

embed_png <- function(path, alt) {
  sprintf("![%s](%s)", alt, knitr::image_uri(normalizePath(path, mustWork = TRUE)))
}
```

## Purpose

This report treats each matrix cell as a subject and analyzes three endpoints:
first infection, first recovery, and death. Cells without a given event are
censored at the final observed simulation step.

The event columns follow the standard `R` survival convention used by
`Surv(time, status)`: `TRUE` means the event was observed, and `FALSE` means the
cell was censored for that endpoint.

```{r load-endpoints}
survival_endpoints <- read_csv(file.path(derived_dir, "survival_endpoints.csv"), show_col_types = FALSE)

survival_endpoints <- survival_endpoints %>%
  mutate(
    initially_infected = as.logical(initially_infected),
    event_infection = as.logical(event_infection),
    event_recovery = as.logical(event_recovery),
    event_death = as.logical(event_death)
  )

head(survival_endpoints)
```

## Event Summary

```{r event-summary}
event_summary <- data.frame(
  endpoint = c("First infection", "Recovery", "Death"),
  events = c(
    sum(survival_endpoints$event_infection),
    sum(survival_endpoints$event_recovery),
    sum(survival_endpoints$event_death)
  ),
  censored = c(
    sum(!survival_endpoints$event_infection),
    sum(!survival_endpoints$event_recovery),
    sum(!survival_endpoints$event_death)
  ),
  median_time = c(
    median(survival_endpoints$time_to_infection),
    median(survival_endpoints$time_to_recovery),
    median(survival_endpoints$time_to_death)
  )
)

event_summary
```

## Kaplan-Meier Curves

```{r km-infection, echo=FALSE, results='asis'}
cat(embed_png(file.path(figure_dir, "km_infection.png"), "Kaplan-Meier infection curve"))
```

```{r km-recovery, echo=FALSE, results='asis'}
cat(embed_png(file.path(figure_dir, "km_recovery.png"), "Kaplan-Meier recovery curve"))
```

```{r km-death, echo=FALSE, results='asis'}
cat(embed_png(file.path(figure_dir, "km_death.png"), "Kaplan-Meier death curve"))
```

## Stratified Kaplan-Meier Curves

Following the reference lab workflow, the infection endpoint is stratified by
whether the cell was infected at step `0`, and the death endpoint is stratified
by distance from the center seed.

```{r stratified-km-infection, echo=FALSE, results='asis'}
cat(embed_png(file.path(figure_dir, "km_infection_initial_status.png"), "Stratified infection curve"))
```

```{r stratified-km-death, echo=FALSE, results='asis'}
cat(embed_png(file.path(figure_dir, "km_death_distance_group.png"), "Stratified death curve"))
```

## Kaplan-Meier Tables

```{r km-tables}
infection_fit <- survfit(Surv(time_to_infection, event_infection) ~ 1, data = survival_endpoints)
recovery_fit <- survfit(Surv(time_to_recovery, event_recovery) ~ 1, data = survival_endpoints)
death_fit <- survfit(Surv(time_to_death, event_death) ~ 1, data = survival_endpoints)

list(
  infection = summary(infection_fit)$table,
  recovery = summary(recovery_fit)$table,
  death = summary(death_fit)$table
)
```

## Log-Rank Tests

Log-rank tests compare survival curves across the same strata used above.

```{r log-rank-tests}
infection_logrank <- survdiff(Surv(time_to_infection, event_infection) ~ initially_infected, data = survival_endpoints)
death_logrank <- survdiff(Surv(time_to_death, event_death) ~ distance_group, data = survival_endpoints)

list(
  infection_by_initial_status = infection_logrank,
  death_by_distance_group = death_logrank
)
```

## Cox Models

The models use row, column, distance from center, and initial infection status as
cell-level covariates. Coefficients are descriptive for this single simulated
run and should not be generalized beyond this baseline scenario.

```{r cox-helper}
fit_cox <- function(time_column, event_column) {
  event_count <- sum(survival_endpoints[[event_column]])
  if (event_count < 2) {
    return(data.frame(
      term = "not_fit",
      estimate = NA_real_,
      conf.low = NA_real_,
      conf.high = NA_real_,
      p.value = NA_real_,
      note = paste("Only", event_count, "event(s); Cox model not fit.")
    ))
  }

  formula <- as.formula(
    paste0("Surv(", time_column, ", ", event_column, ") ~ row + col + distance_from_center + initially_infected")
  )

  broom::tidy(coxph(formula, data = survival_endpoints), exponentiate = TRUE, conf.int = TRUE) %>%
    mutate(note = "Hazard ratio")
}
```

```{r cox-results}
cox_results <- bind_rows(
  fit_cox("time_to_infection", "event_infection") %>% mutate(endpoint = "First infection"),
  fit_cox("time_to_recovery", "event_recovery") %>% mutate(endpoint = "Recovery"),
  fit_cox("time_to_death", "event_death") %>% mutate(endpoint = "Death")
) %>%
  select(endpoint, term, estimate, conf.low, conf.high, p.value, note)

cox_results
```

## Interpretation

```{r interpretation}
interpretation <- data.frame(
  finding = c(
    "Cells are censored independently for each endpoint.",
    "Initially infected cells have infection time zero.",
    "Death is only possible because mortality is enabled in the baseline run.",
    "Cox model estimates are descriptive because this report analyzes one simulated baseline run."
  )
)

interpretation
```
